---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

by Gabriel Espinosa

## Loading and preprocessing the data

First, load the data in the environment. The code changes the column date from char to Date type.

```{r loadData}
  library(plyr)
  activity <- read.csv(unzip("activity.zip"))
  activity$date <- as.Date(activity$date)
  str(activity)
```



## What is mean total number of steps taken per day?

First, the total Steps per day are reported.

```{r sumSteps}

```


The next code generates the **histogram** which shows the total Steps per day.

```{r histSteps}

```


Now, it calculates the **mean** and **median** of the total Steps per day.

```{r summarySteps}

```



## What is the average daily activity pattern?

Here, the code draws the plot of 5-min interval steps taken (x-axis) and average of the steps taken across all days.

```{r linePDailyPattern}

```


For the second part of the question, you can see the red point drawn by the next code, which indicates the 5-minutes interval with maximum (average) number of steps across all days.

```{r pointMaxDailyPattern}

```



## Imputing missing values

First, the code reports the total number of missing values (**NA**) in the data set.

```{r totalNA}

```

For the second part, I decided to use the mean of the 5-min interval across all days for imputing the data.  

Now, it creates a new data set with the imputed data, as specify before.

```{r newImputData}

```


It's time to draw an histogram of the new (**imput**) data set.

```{r histImputData}

```


As it shows in this results, it differs from ...



## Are there differences in activity patterns between weekdays and weekends?

For the last question, the next code creates a new column (variable) with two factors: *weekday* or *weekend*, this indicates that the date is in a "weekday" day or "weekend" day. 

```{r createFactorData}
library(lubridate)
# wday(Sys.Date(), week_start = TRUE)
```


Finally, a line plot with the two factors, wih 5-min interval steps taken (x-axis) versus average of the steps taken across all days, of course, for each factor (*weekday/weekend days*).

```{r plotWeekSteps}

```

Thanks for your attention.

