---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

by Gabriel Espinosa

## Loading and preprocessing the data

First, load the libraries and data in the environment. The code changes the column date from char to Date type.

```{r loadLibraries, message=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r loadData, cache=TRUE}
activity <- read.csv(unzip("activity.zip"))
activity$date <- as.Date(activity$date)
str(activity)
```
  


## What is mean total number of steps taken per day?

First, the total Steps per day are reported. Then, it generates the **histogram** which shows the total Steps per day.

```{r sumSteps, warning=FALSE}
with(activity,tapply(steps, date, sum, na.rm = TRUE))

actDay <- activity %>% group_by(date) %>% summarise(totalsteps = sum(steps,na.rm = TRUE))
actDay$date <- as.Date(actDay$date)
ggplot(actDay,aes(date,totalsteps)) + geom_histogram(stat="identity")  + xlab("Date") + ylab("Total Steps") + ggtitle("Total Steps per day")
```


  
Now, it calculates the **mean** and **median** of the total Steps per day.

```{r summarySteps}
activity %>% group_by(date) %>% summarise(meanSteps = mean(steps,na.rm = TRUE), medianSteps = median(steps, na.rm = TRUE))
```

  

## What is the average daily activity pattern?

Here, the code draws the plot of 5-min interval steps taken (x-axis) and average of the steps taken across all days.  
For the second part of the question, you can see the red point which indicates the 5-minutes interval with maximum (average) number of steps across all days.

```{r linePDailyPattern}
actInt <- activity %>% group_by(interval) %>% summarise(averageSteps = mean(steps,na.rm = TRUE))

with(actInt,plot(interval, averageSteps, type="l", main = "Average steps for 5-minutes interval across all days", xlab = "5-minute interval", ylab = "Average Steps", xaxt="n"))
# Just adding format to x-axis.
t <- format( seq.POSIXt(as.POSIXct(Sys.Date()), as.POSIXct(Sys.Date()+1), by = "5 min"),"%H:%M", tz="GMT")
axis(1,seq(0,2355,by = 105),t[seq(1,288,by=13)],las=2)

with(actInt,points(interval[which.max(averageSteps)],max(averageSteps), col="red", pch = 19))
# Max average steps: 
with(actInt,interval[which.max(averageSteps)])
```


  
## Imputing missing values

First, the code reports the total number of missing values (**NA**) in the data set. We can see in the result, the 2304 **NA** are in the *steps* column.

```{r totalNA}
colSums(is.na(activity))
```

For the second part, I decided to use the mean of the 5-min interval across all days (*actInt* data set) for imputing the data.  

Now, it creates a new data set with the imputed data, as specify before.

```{r newImputData}
actImput <- activity
actImput$steps <- replace(actImput$steps,is.na(actImput$steps),actInt$averageSteps)
actImput$steps <- sapply(actImput$steps,round)
str(actImput)
```


It's time to draw an histogram of the new (**imput**) data set.

```{r histImputData, warning=FALSE}
actImputDay <- actImput %>% group_by(date) %>% summarise(totalSteps = sum(steps,na.rm = TRUE))
actImputDay$date <- as.Date(actImputDay$date)
ggplot(actImputDay,aes(date,totalSteps)) + 
  geom_histogram(stat="identity") + 
  xlab("Date") + ylab("Total Steps") + 
  ggtitle("Total Steps per day (Imputed data)")

actImput %>% group_by(date) %>% summarise(meanSteps = mean(steps), medianSteps = median(steps))
```


As it shows in this results, the results really change for the days with **NaN mean**, thats because all the intervals that day are **NA**. With this imputing "technic", it's obviuos because I decided to round the values to a integer, i.e. the NA's in interval 0 (*00:00*) change to 2 steps, rounded from 1.7169.


  
## Are there differences in activity patterns between weekdays and weekends?

For the last question, the next code creates a new column (variable) with two factors: *weekday* or *weekend*, this indicates that the date is in a "weekday" day or "weekend" day. 

```{r createFactorData}
actWday <- actImput %>% mutate(weekday = ifelse(wday(date, week_start = TRUE) %in% 1:5, "weekday","weekend"))
actWday$weekday <- as.factor(actWday$weekday)
str(actWday)
```


Finally, a line plot with the two factors, wih 5-min interval steps taken (x-axis) versus average of the steps taken across all days, of course, for each factor (*weekday/weekend days*).

```{r plotWeekSteps, warning=FALSE}
actWdayInt <- actWday %>% group_by(interval, weekday) %>% summarise(averageSteps = mean(steps,na.rm = TRUE))

ggplot(actWdayInt,aes(interval,averageSteps, color=as.factor(weekday))) + 
    facet_grid(weekday~.) + 
    geom_line() + 
    xlab("Average Steps") + 
    ylab("5-minutes interval") + 
    ggtitle("Average steps for 5-minutes interval on weekdays and weekend days") +
    theme(legend.position = "none")
```
  
### Thanks for your attention.

